---
title: "Use of Pipe oprator"
author: "Sadaf Zuhra"
date: "2024-09-17"
output: 
  pdf_document: 
    highlight: zenburn
  html_document: 
    highlight: zenburn
    theme: journal
---

# Combining Multiple Operations with the Pipe
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(nycflights13)
by_dest <- group_by(flights, dest)
delay <- summarize(by_dest,
 count = n(),
  dist = mean(distance, na.rm = TRUE),
 delay = mean(arr_delay, na.rm = TRUE))

delay <- filter(delay, count > 20, dest != "HNL")

```
There’s another way to tackle the same problem with the pipe, %>%:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
delays <- flights %>%
 group_by(dest) %>%
 summarize(
 count = n(),
 dist = mean(distance, na.rm = TRUE),
 delay = mean(arr_delay, na.rm = TRUE)
 ) %>%
 filter(count >  20, dest != "HNL")

```

This focuses on the transformations, not what’s being transformed,
which makes the code easier to read. You can read it as a series of
imperative statements: group, then summarize, then filter. As sug‐
gested by this reading, a good way to pronounce %>% when reading
code is “then".
# Missing Values
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
flights %>%
 group_by(year, month, day) %>%
 summarize(mean = mean(dep_delay))

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
flights %>%
 group_by(year, month, day) %>%
 summarize(mean = mean(dep_delay, na.rm = TRUE))
 
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
not_cancelled <- flights %>%
 filter(!is.na(dep_delay), !is.na(arr_delay))

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
not_cancelled %>%
 group_by(year, month, day) %>%
 summarize(mean = mean(dep_delay))

```












